The fundamental shape of dynamic experiences is preserved in episodic memory

### 2018-07-09 main.tex
added methods up to list learning analyses

>>>>

\section{Methods}
\label{sec:methods}

\subsection{Participants and Experimental Design}
How much detail here? or just point to janice's paper?

\subsubsection{Fitting the topic model to the video text and recall transcripts}
A topic model was used to estimate the most likely mixture of topics for a given sample of text. First, the video was manually segmented into 1000 scenes, and a collection of descriptive features was manually transcribed. For each scene, we considered the following features: scene details (i.e. a sentence or two describing what happened in that scene, space (indoor or outdoor), name of all the characters in the scene, name of the character in focus, name of the character speaking, location, camera angle, music presence, and words on the screen. We concatenated the text for all of these features within each segment, creating a 'bag of words' describing each scene. We then transformed the text descriptions into overlapping windows of 50 scene segments. For example, the first text sample comprised the text from the first 50 segments, the second comprised the text from n+1:n+51, and so on. We trained our model using these overlapping text samples using scikit-learn's (version 0.19.1) `CountVectorizer' and `LatentDirichletAllocation' classes.  First, the text was transformed into a vector of word counts (default parameters). This gave a word count vector for each scene in the video.  Then, the word count vectors were used to fit a topic model (\# of topics=100, method=batch). We transformed the text descriptions using the model resulting in a scenes (1000) by topics (100) matrix. The scene descriptions often spanned multiple timepoints (i.e. TRs). To account for this, we expanded the video model by copying the rows of the model for as many timepoints that the scene description spanned. After this expansion, the shape of the model was the length of the duration of the video (1976 TRs).

To create the recall models, for each participant we tokenized the recall transcript into a list of sentences and then mapped the list to overlapping windows of 10 sentences.  We transformed the list of overlapping recall sentences using the model that was trained on the video text (as described in the paragraph above). The result of this was a sentences (range: 68-294) by topics (100) matrix for each participant that represented the most likely mixture of topics for a given chunk of sentences.

\subsubsection{Extracting events using a hidden Markov model}
The topic model timepoint-by-timepoint correlation matrices all exhibited a block-diagonal structure (with small off-diagonal values), suggesting that the models were comprised of a number of sequential `states' (or events). To capture this structure, we fit the video and each recall model using a hidden Markov model (HMM). Given a number of states or events ($k$), the HMM recovers a set of labels that segments consecutive timepoints into $k$ events (REF markov paper). To implement this analysis, we used the Brainiak toolbox \citep{BaldEtal17}.

Our metric for choosing the `best fitting' HMM was to choose the model with the $k$ value that maximized the ratio of the average `within-event' correlation values (i.e. the correlation values for blocks of consecutive timepoints the model identified as one event) and the average `across-event' correlation (i.e. the rest of the correlation values). Additionally, we included a penalty parameter that was proportional to the smoothing of the model that preferred models with smaller $k$ values. We chose $k$ values separately for the video model and for each recall model.  Then, using the best $k$ values, We fit a separate HMM for each topic model. Finally, we averaged over timepoints identified to be in the same event resulting in a $k$ events by 100 topics matrix for the video model and each of the recall models.
%
% \subsubsection{Matching recall events to video events}
% To figure out which video event each recall event referred to, we correlated the video events model and each recall events model. This resulted in a video events (34) by recall events (8-27) correlation matrix (for each participant) which contains the similarity between each video event and each recall event.  To find the most likely video event that a given recall event referred to, we computed the argmax over the columns of this matrix.  The result was a list of video event indices for each participant. These indices are analogous to the values found in a ``recall matrix'' from a free recall list learning experiment, but represent the recall of particular events (instead of words, for example).
%
% \subsubsection{Characterizing memory performance}
% We used a few different metrics to assess memory that were all designed to capture the relationship between the video and recall models.  First, to get a general sense for the match between the video and recall matrices, we vectorized the the matrices and correlated them.  This resulted in a single number representing how well a given participant recalled the video. To get a temporally dynamic measure of memory, for each participant we computed the pairwise correlation between matching video/recall timepoints.  This allowed us to assess memory for individual moments of the video. Finally to get a more nuanced representation of memory for the video, we computed a video/recall model correlation matrix for each participant. These timepoint (1976) by timepoint (1976) matrices represent the correlation between every moment of the video model and every moment of the recall model. To provide some intuition, if a participant recalled every scene at the same rate and in exactly the same order as the video, we would expect high correlation values along the diagonal of this matrix. However, if a participant recalls the scenes out of order, or at different rates for different scenes, this will result in high values off the diagonal.
%
% The only additional constraint we imposed was that there was at least 2 points between each line segment.
% We searched for the optimal number of line segments by iteratively splitting the participants into two groups, fitting the model to the first half and testing the model on the remaining subjects. To compute fit to the model, for each recall event (a 1 by 100 topics vector) we computed the nearest line segment and then calculated the mean squared distance averaging over all events and all participants (in the sample). We repeated this 20 times using 1-10
%
% \subsubsection{Bridging between the stimulus and recall}
% The participants in this study were instructed to recall as much of the episode as they could in the order it happened.  Thus, we hypothesized that the recall models should resemble the video model to the extent that participants followed these instructions and had reasonably good memory accuracy. To test whether this was true, we correlated each moment of the video model with each moment of each participants' resampled recall model. We then averaged these matrices together to get a single group averaged matrix.  The result was a viewing time (1976) by recall time (1976) correlation matrix representing the average relationship between every moment of the movie and recall models (Fig.~\ref{fig:model}e). Notably, we observe significant correlation values (p<.05, permutation test described in \ref{sec:methods}) primarily along the diagonal, suggesting that our model captures the fact that participants recalled the episode in order. We also visualized this effect by projecting the matrices into a 3D space (reduced using Spectral Embedding [cite]) where the proximity of the trajectories at every timepoint represents how similar the topic vectors were. Visualizing the data in this way makes it clear that the overall shape and correlational structure of the video and average recall matrices is quite similar.
